---
title: "Homework 2"
author: "Megan Sporre"
date: "January 16, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##R Intermediate

# Iris Loops
1. Describe what values are in the loop output.
  The loop is outputing average values of sepal length and width and petal length and width for each species of iris.

2. Describe using pseudo code how the output was calculated.
  
```{r}
data(iris)
sp_ids = unique(iris$Species)
#returns vector of unrepeated species names (class = factor)
output = matrix(0, nrow=length(sp_ids), ncol=ncol(iris)-1)
#output is a matrix where rows are each species and columns are every column but the last column (species). 0 stops the matirx from populating with values (waiting for function to populate)
rownames(output) = sp_ids
#naming rows of output matrix each species from sp_ids (aka pecies names from iris)
colnames(output) = names(iris[ , -ncol(iris)])
#naming columns of output each column name from iris. returns columns 1:4, and not column 5
#below for any species in sp_ids
for(i in seq_along(sp_ids)) {
    iris_sp = subset(iris, subset=Species == sp_ids[i], select=-Species)
    #subset data in iris to be values for one species
    for(j in 1:(ncol(iris_sp))) {
      #created a vector of 1:4 (number of columns in iris_sp)
      #create two objects, x and y, with values equal to 0
        x = 0
        y = 0
        if (nrow(iris_sp) > 0)
          #if the number of rows in iris_sp is greater than 0, perform the following
          {
          #create a vector 1:50 (# of rows in iris_sp)  
          for(k in 1:nrow(iris_sp)) {
                x = x + iris_sp[k, j]
                #defined as the sum of all numbers in iris_sp
                y = y + 1
                #defined as length of iris_sp
            }
            output[i, j] = x / y #output sum of column of iris_sp divided by length of column of iris_sp
        }
    }
}
output
```

3. The variables can be renamed as follows; output = Average , y = length of column, x = sum of column.

4. It is possible to accomplish the same task using fewer lines of code.
```{r}
sp_ids = unique(iris$Species)

output = matrix(0, nrow=length(sp_ids), ncol=ncol(iris)-1)
rownames(output) = sp_ids
colnames(output) = names(iris[ , -ncol(iris)])

for(i in seq_along(sp_ids)) {
    iris_sp = subset(iris, subset=Species == sp_ids[i], select=-Species)
    for(j in 1:(ncol(iris_sp))) {
            output[i, j] = sum(iris_sp[,j]) / nrow(iris_sp) 
        }
    }
output
```

5.Write a for loop that will produce a vector y that contains the sum of x up to that index of x. So for example the elements of x are 1, 2, 3, and so on and the elements of y would be 1, 3, 6, and so on.

```{r}
vector.y=NULL
for(i in 1:10) {
  vector.y[i]=sum(1:i)
}
vector.y
```

6. Modify your for loop so that if the sum is greater than 10 the value of y is set to NA

```{r}
vector.y2=NULL
for(i in 1:10) {
  vector.y2[i]=sum(1:i)
  if(vector.y2[i]<=10){
    print(vector.y2[i])
  }
 else if(vector.y2[i]>10) {
    print('NA')
 }

}

```

7. Place your for loop into a function that accepts as its argument any vector of arbitrary length and it will return y.

```{r}
function.x <-function(n)
{
  vector.y3=NULL
  for(i in 1:n) {
  vector.y3[i]=sum(1:i)
  }
  return(vector.y3)
}
function.x(100)
```

